\thetitle

### 2020-12-04 sun_phases/spin_qudit_tomo.tex
fixed variety of errors involving rotation matrices vs. their conjugates

The polarization operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$ under 3D rotations (see Appendix \ref{sec:rotations}).
For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
\begin{align}
  R\p{\v\omega} \equiv e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
\end{align}
and expand polarization operators in a rotated frame as
\begin{align}
  T_{\v\omega\ell m} \equiv
  R\p{\v\omega}^\dag T_{\ell m} R\p{\v\omega}
  = \sum_{n=-\ell}^\ell D_{nm}^\ell\p{\v\omega} T_{\ell n},
  \label{eq:trans_rot}
\end{align}
where
\begin{align}
  D_{mn}^\ell\p{\v\omega} \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
\end{align}
are (Wigner) rotation matrix elements.
Throughout this work, we will primarily consider rotations of the sphere that take the north pole to a point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
For ease of notation, we will therefore implicitly associate Euler angle doublets $\v v=\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, such that $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{\alpha,\beta,0}$, and $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$.

The polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond their indexing and the rules for their transformation under 3D rotations.
In fact, the phase-space representation of $T_{\ell m}$ is proportional to $Y_{\ell m}$.
The phase-space representation of a spin qudit operator $X$ assigns, to each point $\v v$ on the sphere, the complex number
\begin{align}
  X^{\t{PS}}\p{\v v} \equiv \bk{s_{\v v}|X|s_{\v v}},
\end{align}
where $\ket{s_{\v v}}\equiv R\p{\v v}\ket{s}$ is the state of a spin qudit polarized along $\v v$.
This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $X^{\t{PS}}\p{\v v}$ at all points $\v v$ on the sphere.
The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that (see Appendix \ref{sec:rotations})
\begin{align}
  T_{\ell m}^{\t{PS}}\p{\v v} = c_\ell Y_{\ell m}\p{\v v},
\end{align}
where the scalar $c_\ell$ simply enforces $\obk{T_{\ell m}|T_{\ell m}}=1$.
The polarization operators $T_{\ell m}$ are thus a quantum analogue of the spherical harmonics $Y_{\ell m}$, and play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.

@@

In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\v v\ell,0}}_\rho$, we essentially need to to find a set of coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
\begin{align}
  \rho_{\ell m}^* = \bk{T_{\ell m}}_\rho
  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
  \label{eq:state_recon}
\end{align}
Expanding the rotated polarization operators $T_{\v vk,0}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
\begin{align}
  T_{\ell m}
  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} D^k_{n,0}\p{\v v} T_{kn}.
\end{align}
Orthogonality of the polarization operators then implies the decomposition $C_{\ell mk}\p{\v v}=\delta_{\ell k}C_{\ell m}\p{\v v}$, and in turn
\begin{align}
  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v}
  = \delta_{mn}
  \label{eq:tomo_recovery}
\end{align}
for all $\ell$.
In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
\begin{align}
  f\p{\v v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\v v},
\end{align}
where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v} f_{kn}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
\begin{align}
  \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
  = \delta_{\ell k} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).

If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points on the sphere \cite{freeden2008spherical, freeden2018spherical}.
The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially band-limited at degree $d$ \footnote{Specifically, the phase-space representation of a $d$-level spin qudit state $\rho$ is band-limited at degree $d$.}.
Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f_\ell = \sum_m f_{\ell m} Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are polarization operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.

@@

  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell}.

@@

Here $\v v$ and $\ell$ label a column of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space.

@@

Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv \sum_j x^V_{kj} \ket{j}$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators

@@

  Q^V_k \equiv \sum_j q^V_{kj} T_j,

@@

  \rho_k^V \equiv \bk{Q^V_k}_\rho.

@@

where

@@

  = \bk{Q^V_k}_\rho

@@

  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,

@@

where $M_V^{-1}$ is the right inverse of $M_V$, satisfying $M_V M_V^{-1}=1$.

@@

Computing the classical error scale $\S_V$ and estimates $\tilde\rho^V_k\approx\rho^V_k$ requires building the measurement matrix $M_V$ and computing its singular value decomposition.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a polarization operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$-sized blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
  \label{eq:block}
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular value decomposition takes at most $O(\abs{V}d^2)$ time.
If we assume that $\abs{V}\sim d$, then computing the classical error scale $\S_V$ takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

@@

We now define $\N_{V\ell}$ and $\ket{\chi_{V\ell}}$ for the sake of completion, but note that these definitions can be skipped without consequence for the remaining discussions in this paper.
The matrix $\N_{V\ell}$ is defined in terms of the measurement matrix $M_{V\ell}$ and its right inverse $M_{V\ell}^{-1}$ by
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero.
The vector $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ is defined by
\begin{align}
  \chi^V_{LM} &\equiv
  \sum_\ell \obk{\N_{V\ell}|\D_M|\tilde g_{L\ell}}, \\
  \tilde g_{L\ell} &\equiv \sum_{m,m'}
  \obk{T_{L,m+m'}|T_{\ell m}^\dag T_{\ell m'}} \op{m}{m'}, \\
  \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
\end{align}
Here $\tilde g_{L\ell}$ is essentially a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
\begin{enumerate}
\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
  If $\abs{V}$ is too large for such optimization, you can simply generate many sets of random measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
  Note that computing the error scale $\epsilon_V$ requires, for each $\ell\in\set{0,1,\cdots,d-1}$, constructing the measurement matrix $M_{V\ell}$ in Eq.~\eqref{eq:block} and computing its singular value decomposition.
  Save all measurement matrix data associated with the final measurement axes $V$ for later use. % todo: cite codes
\item For each axis $\v v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
\item Use the the estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ to compute the estimates
\begin{align}
  \tilde T_{\v v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\v v\mu}
\end{align}
of $\bk{T_{\v v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding right singular vectors by $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, compute the coefficients and operators
  \begin{align}
    \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
    x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0}, \\
    Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
    x^V_{\ell k\v v} T_{\v v\ell,0},
  \end{align}
  and combine them into the estimate
  \begin{align}
    \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
    \approx \rho.
  \end{align}
\end{enumerate}
If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$, which can be computed from Eq.~\eqref{eq:error}. % todo: cite codes
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

>>>>

The polarization operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform similarly to spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$ under 3D rotations (see Appendix \ref{sec:rotations}).
Specifically, for any triplet of angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
\begin{align}
  R\p{\v\omega} \equiv e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
\end{align}
and expand rotated polarization operators as
\begin{align}
  T_{\v\omega\ell m} \equiv
  R\p{\v\omega} T_{\ell m} R\p{\v\omega}^\dag
  = \sum_{n=-\ell}^\ell D_{mn}^\ell\p{\bar{\v\omega}}^* T_{\ell n},
  \label{eq:trans_rot}
\end{align}
where $\bar{\v\omega}=\p{\gamma,\beta,\alpha}$ is the reversal of $\v\omega$, and
\begin{align}
  D_{mn}^\ell\p{\bar{\v\omega}}
  \equiv \bk{\ell m|R\p{\bar{\v\omega}}|\ell n}
\end{align}
are (Wigner) rotation matrix elements.
Throughout this work, we will primarily consider rotations of the sphere that take the north pole to a point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
For ease of notation, we therefore define $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$, and $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{0,\beta,\alpha}$.

The polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond the rules for their transformation under 3D rotations.
In fact, the phase-space representation of $T_{\ell m}$ is proportional to $Y_{\ell m}$.
The phase-space representation of a spin qudit operator $X$ assigns, to each point $\v v$ on the sphere, the complex number
\begin{align}
  X^{\t{PS}}\p{\v v} \equiv \bk{s_{\v v}|X|s_{\v v}},
\end{align}
where $\ket{s_{\v v}}\equiv R\p{\v v}\ket{s}$ is the state of a spin qudit polarized along $\v v$.
This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $X^{\t{PS}}\p{\v v}$ at all points $\v v$ on the sphere.
The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that (see Appendix \ref{sec:rotations})
\begin{align}
  T_{\ell m}^{\t{PS}}\p{\v v} = c_\ell Y_{\ell m}\p{\v v},
\end{align}
where the scalar $c_\ell$ simply enforces $\obk{T_{\ell m}|T_{\ell m}}=1$.
The polarization operators $T_{\ell m}$ are thus a quantum analogue of the spherical harmonics $Y_{\ell m}$, and play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.

@@

In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\v v\ell,0}}_\rho$, we essentially need to to find a set of coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
\begin{align}
  \rho_{\ell m}^* = \bk{T_{\ell m}}_\rho
  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
  \label{eq:state_recon}
\end{align}
Expanding the rotated polarization operators $T_{\v vk,0}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
\begin{align}
  T_{\ell m}
  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} D^k_{0,n}\p{\v v}^* T_{kn}.
\end{align}
Orthogonality of the polarization operators then implies the decomposition $C_{\ell mk}\p{\v v}=\delta_{\ell k}C_{\ell m}\p{\v v}$, and in turn
\begin{align}
  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{0,n}\p{\v v}^*
  = \delta_{mn}
  \label{eq:tomo_recovery}
\end{align}
for all $\ell$.

In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
\begin{align}
  f\p{\v v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\v v},
\end{align}
where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v} f_{kn}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
\begin{align}
  \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
  = \delta_{\ell k} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
Due to the fact that $D^\ell_{0,m}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{\ell m}\p{\v v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).

If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points on the sphere \cite{freeden2008spherical, freeden2018spherical}.
The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states (or rather, their phase-space representations) are band-limited at degree $d$.
Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f_\ell = \sum_m f_{\ell m} Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are polarization operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.

@@

  M_V \equiv \sum_{\v v,\ell} \ket{\v v\ell} \obra{T_{\v v\ell,0}}.

@@

Here $\v v$ and $\ell$ label a row of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space, and $\obra{T_{\v v\ell,0}}$ is the conjugate transpose of $\oket{T_{\v v\ell,0}}$.

@@

Indexing these singular values $M^V_k$ and the corresponding (normalized) left singular vectors $\v x^V_k\equiv \sum_j x^V_{kj} \ket{j}$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators

@@

  Q^V_k \equiv \sum_j \p{q^V_{kj}}^* T_j,

@@

  \rho_k^V \equiv \bk{{Q^V_k}^\dag}_\rho.

@@

where, using the fact that $T_j=T_j^\dag$,

@@

  = \bk{{Q^V_k}^\dag}_\rho

@@

  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'} \,

@@

where $M_V^{-1}$ is the left inverse of $M_V$, satisfying $M_V^{-1} M_V = 1$.

@@

Computing the classical error scale $\S_V$ and estimates $\tilde\rho^V_k\approx\rho^V_k$ requires building the measurement matrix $M_V$ and computing its singular value decomposition.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a polarization operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  M_V U = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\abs{V}\times\p{2\ell+1}$-sized blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{\v v,m} \ket{\v v}
  \obk{T_{\v v\ell,0}|T_{\ell m}} \bra{m}
  = \sum_{\v v,m} D^\ell_{0,m}\p{\v v} \op{\v v}{m}.
  \label{eq:block}
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the left inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular value decomposition takes at most $O(\abs{V}d^2)$ time.
If we assume that $\abs{V}\sim d$, then computing the classical error scale $\S_V$ takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

@@

We now define $\N_{V\ell}$ and $\ket{\chi_{V\ell}}$ for the sake of completion, but note that these definitions can be skipped without consequence for the remaining discussions in this paper.
The matrix $\N_{V\ell}$ is defined in terms of the measurement matrix $M_{V\ell}$ and its left inverse $M_{V\ell}^{-1}$ by
\begin{align}
  \N_{V\ell} \equiv M_{V\ell}^\dag
  \diag\sp{\p{M_{V\ell}^{-1}}^\dag M_{V\ell}^{-1}} M_{V\ell},
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero.
The vector $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ is defined by
\begin{align}
  \chi^V_{LM} &\equiv
  \sum_\ell \obk{\N_{V\ell}|\D_M|\tilde g_{L\ell}}, \\
  \tilde g_{L\ell} &\equiv \sum_{m,m'}
  \obk{T_{L,m+m'}|T_{\ell m}^\dag T_{\ell m'}} \op{m}{m'}, \\
  \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
\end{align}
Here $\tilde g_{L\ell}$ is essentially a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
\begin{enumerate}
\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
  If $\abs{V}$ is too large for such optimization, you can simply generate many sets of random measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
  Note that computing the error scale $\epsilon_V$ requires, for each $\ell\in\set{0,1,\cdots,d-1}$, constructing the measurement matrix $M_{V\ell}$ in Eq.~\eqref{eq:block} and computing its singular value decomposition.
  Save all measurement matrix data associated with the final measurement axes $V$ for later use. % todo: cite codes
\item For each axis $\v v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
\item Use the the estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ to compute the estimates
\begin{align}
  \tilde T_{\v v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\v v\mu}
\end{align}
of $\bk{T_{\v v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding left singular vectors by $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, compute the  operators and coefficients
  \begin{align}
    Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
    \p{x^V_{\ell k\v v}}^* T_{\v v\ell,0}, \\
    \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
    x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0},
  \end{align}
  and combine them into the estimate
  \begin{align}
    \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
    \approx \rho.
  \end{align}
\end{enumerate}
If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$, which can be computed from Eq.~\eqref{eq:error}. % todo: cite codes
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
